Watch Out for the Bully! Job Interference Study on Dragonfly Network

### 2016-03-29 draft.tex
tweak some sentences, change some words

The low latency and high bandwidth interconnect network plays a critical role in HPC system performance. As computational capabilities continue to grow due to many-core nodes with accelerators, the network is increasingly becoming a scarce resource. The high-radix, low-diameter dragonfly topology can lower the \NOTE{?} cost of the interconnect, improve network bandwidth and reduce packet latency \cite{dally-dragonfly}. This make the dragonfly network a very promising choice for building petaflop and beyond supercomputers. 
%The procurement results from Trinity and CORAL project shows the great performance improvement can be got from using dragonfly network.

@@

In this work, we study the interference between concurrently running jobs on dragonfly networks with different job placement and routing policies. We use CODES, an HPC network simulation toolkit with high fidelity and excellent scalability for packet level simulation \cite{codes}. Unlike much existing work that rely on synthetic workloads, we use traces of three representative applications from the DOE Design Forward Project \cite{designforwardwebpage}. The advantage of using real application traces is straight forward: \TODO{these traces can capture detailed application communication behaviors that are  omitted in synthetic workload for simplification}. Applications are usually distinctive from each other in data transfer amount, communication pattern, operation dependency etc., which can be impractical to reproduce in a synthetic workload.

@@

\subsection{Dragonfly Network}
\label{sec:network}
The dragonfly is a two-level hierarchical topology. It consists of several groups connected by all-to-all links. In each group, there could be ``a" number of routers connected via all-to-all local channels. For each router, there could be ``p" number of compute nodes attached to it via terminal links. The router has ``h" number of global channels that used as intergroup connections through which the router connected to other routers in different groups. Thus, the radix of each router is $k = a+h+p-1$. The implementation of such dragonfly network could choose different value for ``a", ``p" and ``h" in different computing centers. The adoption of proper $a$, $h$, $p$ could be involving many factors, such as system scale, building cost and workload requirements.

For the purpose of load balance, it is recommended that the proper dragonfly configuration should follow $a=2p=2h$\cite{kim-micro}. Under this configuration, the total number of groups ``g" in the dragonfly network would be $g = a*h+1 $, the total number of compute nodes ``N" in the network would be $N = p*a*g $. An implementation of dragonfly network follows such configuration is illustrated in Figure \ref{fig:dragonfly-overview}. There are 6 routers in each group, each router has 3 compute nodes attached to it. This dragonfly implementation consists of 19 groups and 342 nodes in total.

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.48\textwidth]{dragonfly-overview}
  \caption{Dragonfly Network Overview. Illustration of part of dragonfly network. There are 19 groups, 6 routers in each group, 3 compute nodes connected to each router, 342 nodes in total. Job $J1$ requires for 7 nodes, gets its allocation by random placement. Job $J2$ requires for 8 nodes, gets its allocation by contiguous placement. }
  \label{fig:dragonfly-overview}
\end{figure}

@@

\textbf{Contiguous Placement:} In this policy, the compute nodes are assigned to job in a consecutive manner. The assignment fills up a group first, only crosses group boundaries and start to fill the next group as necessary. Job $J2$, illustrated as blue in Figure \ref{fig:dragonfly-overview}, gets the required amount of nodes by contiguous placement. Contiguous placement confines the tasks of a job into the same group, which may result in local network congestion and increase the possibility of hot-spots.

\textbf{Random Placement:} In this policy, job will get a set of nodes that randomly selected from all the available nodes in the system. The nodes assigned to each job would be attached to different routers in different groups, as illustrated in Figure \ref{fig:dragonfly-overview} the red nodes assigned to job $J1$. Some routers may be shared by different jobs when random placement in use. Random placement can distribute the tasks of  a job uniformly across the whole system, thus avoid the possible local congestion and reach load balance. However, random placement may cause possible congestion on global links.

\subsection{Routing on Dragonfly}
\label{sec:routing-schemes}
The routing policy refers to the strategy adopted to route messages \NOTE{(message refers to a stream of packets)} from the source router to the destination router. Previously studied routing policies for dragonfly network include minimal routing, adaptive routing \cite{dally-dragonfly}, and progressive adaptive routing\cite{jiang}, and some of their variations \cite{won-prog-adaptive}. In this work we study three widely adopted routing policies on dragonfly networks.

\textbf{Minimal:} In this policy, a message from source router will take the the shortest path to the destination router. When there are multiple shortest paths between the source and destination router, the message will be evenly divided among the paths. Minimal routing can guarantee the minimum hops the message takes from source to destination. However, it may result in congestion along the shortest path.

\textbf{Adaptive:} In this policy, the path a message takes will be adaptively chosen between shortest and non-shortest path, depends on the congestion situation along that path. For the non-shortest path, an intermediate router will be randomly chosen. The message will first take the shortest path from source to that intermediate router, then it will take the shortest path from that intermediate router to the destination. Since routing path are chosen adaptively based on the congestion situation, it can avoid local congestion and possible hot-spots. Adaptive routing collapses to minimal routing in the absence of traffic congestion.

\textbf{Progressive Adaptive:} In this policy, the decision to route minimally at each hop in the source group will be re-evaluated. And any decision to route non-minimally at source router or at a subsequent hop is permanent and won't be re-evaluated \cite{jiang}. The progressive adaptive is capable of handling the case where a global link is congested but this information has not yet propagated back to the source router \cite{jiang}. The packet will take the non-minimal route only when the congestion is encountered.

@@

It is difficult to do experiment with concurrently running jobs on HPC system\cite{zhou-ipdps-2015}\cite{jain-sc14}\cite{bhatele-sc11}\cite{jokanovic-ipdps-2015}. One reason is that job placement and routing policies are part of system configuration, it is impossible for user to make changes at will. Another reason is that it is unrealistic to reserve the system exclusively to run the same batch of jobs with desired placement and routing configurations and compare the results. Last but not least, configurable dragonfly networks that allow us to perform the exploration presented in this paper are hard to come by for the time being. Therefore, we resort to simulation in our work.

\subsection{Simulation Tool}
\label{sec:simulation-tool}
A simulation toolkit named CODES, Co-Design of Multilayer Exascale Storage Architectures, enables the exploration of simulating different HPC networks with high fidelity and great scalability \cite{codes}. CODES supports dragonfly network\cite{codes-dragonfly} \cite{misbah-tpds}, as well as other topologies such as torus\cite{misbah-pads-2014}\cite{ning-pads-2011} and fat-tree\cite{ning-pads-2015}, with packet-level high-fidelity simulation. It can drive these models through an MPI simulation later that utilizing DUMPI traces\cite{sst}. In this work, we perform an in-depth analysis of job interference with different job placement and routing policies on dragonfly network.

@@

\NOTE{justify network size}The dragonfly network studied in this work consists of 33 groups, each group contains 8 routers and each router has 4 compute nodes attached to it. Overall, there are 264 routers and 1056 compute nodes in the network. The ratio between aggregated local and global channel bandwidth are set propotional to the Cray Cascade system\cite{faanes} based on the dragonfly topology\cite{dally-dragonfly}.

In this work, we study both the network behavior and the job interference when dragonfly network is configured with 6 different job placement and routing policy combinations, which are summarized in Table \ref{tab: placement routing configs}.

@@

    \item \textbf{Network Traffic:} The traffic refers to the amount of data go through each router in the network. For each router, the traffic go through its terminal link, local channel and global channel are analyzed respectively.

    \item \textbf{Network Saturated Time:} The saturated time refers to the time period when  the buffer of certain port in the router is full. Saturated time can denote the congestion on certain port.

@@

\NOTE{limitations:}Since our study is about network behavior and job interference during communication, the time spent by each MPI rank on computation will not be counted in our experiments.

With respect to random placement, we conduct our experiments with 50 sets of distinctive allocation generated by random placement. The corresponding experimental results are median chosen, which intended to eliminate the possibility of variation.

@@

In this study, we use two sets of parallel workloads. Workload \Rmnum{1} consists of AMG, MultiGrid and CrystalRouter. As we see in Table \ref{tab:apps-detail}, AMG has the least amount of data transfer, thus contributes the least amount of traffic in Workload \Rmnum{1}. Workload \Rmnum{2} consists of sAMG, MultiGrid and CrystalRouter. sAMG, a synthetic version of AMG, is generated by increasing AMG data transfer amount by 100x. sAMG contributes the dominant amount of traffic in Workload \Rmnum{2}.

@@

Table \ref{tab:apps-detail} summarizes the details of each application. It shows the number of MPI ranks, the average data transfer amount of each rank and the total data transfer amount for each application.

@@

\section{Study of Parallel Workload \Rmnum{1 }}
%\section{Identify the ``Bully"}
\label{sec:workload-1}

We study the network behavior as well as application interference from two perspectives. First, we analyze the network performance when Workload \Rmnum{1 } is running under different placement and routing configurations. Then, we isolate each application from the workload, and analyze the interference by scrutinizing the traffic go through the routers that serving each application. These analysis allow us to identify the ``bully" in the workload.

@@

The average communication time spent by all MPI ranks when Workload \Rmnum{1} is running under different placement and routing configurations shown in Table \ref{tab:wkld-commtime}. The contiguous placement coupled with minimal routing (CM) results in highest communication time. When contiguous placement coupled with progressive adaptive and adaptive routing (CA, CPA), Workload \Rmnum{1} spends comparable amount of time to complete communication. The random placement policy performs comparably with all routing policies and outperforms all contiguous placement configurations.

@@

    \centering   
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/gc-stime}
        \caption{Global Channel Saturated Time}
        \label{fig:global-channel-stime}
    \end{subfigure}\hfill
     \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/lc-stime}
        \caption{Local Channel Saturated Time}
        \label{fig:local-channel-stime}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/tl-stime}
        \caption{Terminal Link Saturated Time}
        \label{fig:terminal-link-stime}
    \end{subfigure}%
   \caption{The aggregate traffic and saturated time of links in different level in the dragonfly network when Workload \Rmnum{1} running under 6 different placement and routing configurations. ``CA" and ``CPA" have comparably same performance and the corresponding lines are overlapped. }
   \label{fig:wkld-network-traffic-stime}
\end{figure*}

Figure \ref{fig:wkld-network-traffic-stime} shows the aggregate traffic for terminal links, local channels and global channels, and the corresponding saturated time when Workload \Rmnum{1} is running on dragonfly network with 6 different placement and routing configurations.

%global link traffic and busy time explanation
When Workload \Rmnum{1} running with contiguous placement and minimal routing(CM), all its traffic is happening in the contiguously allocated groups, causing congestion along certain shortest paths from source routers to destination routers. Both the traffic through the global channels in those congested paths and the corresponding saturated time are the highest compared with other configurations, shown as the blue dashed line in Figure \ref{fig:global-channel-traffic} and \ref{fig:global-channel-stime}.
When contiguous placement is coupled with (progressive) adaptive routing (CA and CPA), the traffic through congested global channels will be greatly reduced, so does corresponding saturated time, shown as the red and green dashed line in Figure \ref{fig:global-channel-traffic} and \ref{fig:global-channel-stime}. Since (progressive) adaptive routing are congestion aware, the traffic would be redirected to the intermediate routers in other idle groups when congestion occurs, thus the traffic burden on the contiguously allocated groups will be alleviated.

Random placement policy performs comparably with three different routing policies.
Shown as the solid lines in Figure \ref{fig:global-channel-traffic}, the traffic go through global channels are quite balanced across all the routers in the network, thus no particular global channel will suffer high traffic burden. When the random placements coupled with minimal routing (RM), the packets can avoid unnecessary intermediate forwarding as introduced by adaptive and progressive routing, which results in generating less traffic. That's why the blue solid line is lower than the red and green solid line. Since the random placement can uniformly distribute traffic, there are no sudden surge of saturated time of global channel over the network, shown as solid lines in Figure \ref{fig:global-channel-stime}.

@@

We observe similar phenomena from the local channel traffic (Figure \ref{fig:local-channel-traffic}) and saturated time (Figure \ref{fig:local-channel-stime}). Contiguous placement coupled with minimal routing (CM) results in local congestion, and thus high aggregate traffic on the local channels of certain routers. Even with adaptive or progressive adaptive routing, the local congestion caused by grouping all MPI ranks in the same group(s) in contiguous placement can not be eliminated. Random placement is effective at addressing this type of congestion.

%terminal link traffic and busy time explanation
Figure \ref{fig:terminal-link-traffic} is for the purpose of symmetry, it shows the traffic per rank (aka traffic through terminal links), which won't change when different placement and routing configurations in use. All 6 lines are overlapped in Figure \ref{fig:terminal-link-traffic}. However, due to the traffic changes on global and local channels, the saturated time of terminal links will experience variability when different routing schemes in use, as shown in Figure \ref{fig:terminal-link-stime}. When contiguous placement coupled with minimal routing(CM), the accumulated congestion on local and global channels cause extremely long saturated time on the terminal link, shown as the blue dashed line in Figure \ref{fig:terminal-link-stime}. This again manifests the severe congestion caused by using contiguous placement and minimal routing.

@@

As shown in Figure \ref{fig:mg-commtime}, \ref{fig:cr-commtime}, MultiGrid and CrystalRouter suffer the longest communication time when running with contiguous placement and minimal routing (CM). When contiguous placement coupled with adaptive or progressive adaptive routing (CA, CPA), the communication time of both MultiGrid and CrystalRouter is significantly reduced. Random placement further reduces their communication time, and reaches the best performance when coupled with (progressive) adaptive routing (RPA, RA). Both MultiGrid and CrystalRouter show the same trend as the workload when running under different placement and routing configurations.

AMG is quite an exception. As shown in Figure \ref{fig:amg-commtime}, the communication time of AMG skyrockets when running under random placement coupled with (progressive) adaptive routing (RPA, RA). The contiguous placement coupled with (progressive) adaptive routing can reach the best performance for AMG. The performance of AMG running under contiguous placement coupled with minimal routing (CM) is no better than random coupled with minimal routing (RM), but still much better than random with (progressive) adaptive routing (RPA, RA).

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/amg/lc-traffic}
        \caption{AMG Local Channel Traffic}
        \label{fig:amg-lc-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/mg/lc-traffic}
        \caption{MG Local Channel Traffic}
        \label{fig:mg-lc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/cr/lc-traffic}
        \caption{CR Local Channel Traffic}
        \label{fig:cr-lc-traffic}
    \end{subfigure}\\
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/amg/gc-traffic}
        \caption{AMG Global Channel Traffic}
        \label{fig:amg-gc-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/mg/gc-traffic}
        \caption{MG Global Channel Traffic}
        \label{fig:mg-gc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/cr/gc-traffic}
        \caption{CR Global Channel Traffic}
        \label{fig:cr-gc-traffic}
    \end{subfigure}%
   \caption{The aggregate traffic go through the local and global channels of routers serving specific application. ``CA" and ``CPA" have comparably same performance and the corresponding lines are overlapped. More routers are involved in serving each application when random placement is in use, compared with contiguous placement.}
   \label{fig:3app-lc-gc-traffic}
\end{figure*}

@@

\emph{The performance of less communication-intensive jobs in workload will suffer degradation when network random placement and adaptive routing are in use.} AMG in Workload \Rmnum{1} are ``bullied" by their concurrently running communication-intensive peers, MultiGrid and CrystalRouter. AMG shares routers and groups with MultiGrid and CrystalRouter when random placement is in use. The traffic from MultiGrid and CrystalRouter will be redirected to the routers that serving AMG, slow down AMG's communication. \footnote{We have tried three different congestion sensing schemes that used in adaptive routing\cite{won-prog-adaptive}, although there are some variations between the results, none of the congestion sensing scheme can prevent the ``bully" from happening.}

\emph{Consistent performance of each application can be guaranteed when contiguous placement and minimal routing are in use.} The router and group sharing between jobs are prohibited in contiguous placement policy. The traffic won't be redirected from congested routers to less busy ones when minimal routing is in use. Thus the interference between concurrently running jobs will be eliminated and the ``bully" behavior will be terminated. However, contiguous placement and minimal routing usually results in local congestion and unbalanced network utilization.

@@

In this section, we will verify the observations made in the previous section when Workload \Rmnum{1 } is running on dragonfly network under different placement and routing configurations. By generating a synthetic application, we turn the ``bully" into ``bullee".

The synthetic application, sAMG, is generated by increasing AMG data transfer amount by 100x. We run the Workload \Rmnum{2 } that consists of sAMG, MultiGrid and CrystalRouter on the same dragonfly network with different placement and routing configurations. As shown in Table \ref{tab:apps-detail}, sAMG become the most communication-intensive application in Workload \Rmnum{2 }.

@@

The average communication time spent by all MPI ranks when Workload \Rmnum{2} is running under different placement and routing configurations shown in Table \ref{tab:syn-wkld-commtime}. Since the total data transfer amount of Workload \Rmnum{2} increases a lot due to the presence of sAMG, the time spent on communication under all configurations increase as well. The contiguous placement coupled with minimal routing results in highest communication time. The random placement coupled with minimal routing is the most efficient. Progressive adaptive routing is better than adaptive routing when random placement is in use. And they have comparable results when coupled contiguous placement. The performance of Workload \Rmnum{2} under different configurations keep the same trend as it of Workload \Rmnum{1}.

>>>>

The low latency and high bandwidth interconnect network plays a critical role in HPC system performance. As computational capabilities continue to grow due to many-core nodes with accelerators, the network is increasingly becoming a scarce resource. The high-radix, low-diameter dragonfly topology can lower the extension cost of the interconnect, improve network bandwidth and reduce packet latency \cite{dally-dragonfly}. This make the dragonfly network a very promising choice for building petaflop and beyond supercomputers. 
%The procurement results from Trinity and CORAL project shows the great performance improvement can be got from using dragonfly network.

@@

In this work, we study the interference between concurrently running jobs on dragonfly networks with different job placement and routing policies. We use CODES, an HPC network simulation toolkit with high fidelity and excellent scalability for packet level simulation \cite{codes}. Unlike much existing work that rely on synthetic workloads, we use traces of three representative applications from the DOE Design Forward Project \cite{designforwardwebpage}. The advantage of using real application traces is straight forward: these traces can capture detailed application communication behaviors that are  omitted in synthetic workload for simplification. Applications are usually distinctive from each other in data transfer amount, communication pattern, operation dependency etc., which can be impractical to reproduce in a synthetic workload.

@@

\subsection{Dragonfly Network}
\label{sec:network}
The dragonfly is a two-level hierarchical topology. It consists of several groups connected by all-to-all links. In each group, there could be ``a" number of routers connected via all-to-all local channels. For each router, there could be ``p" number of compute nodes attached to it via terminal links. The router has ``h" number of global channels that used as intergroup connections through which the router connected to other routers in different groups. Thus, the radix of each router is $k = a+h+p-1$. Different computing centers could choose different values for ``a", ``p" and ``h" when deploying their dragonfly network. The adoption of proper $a$, $h$, $p$ involves many factors, such as system scale, building cost and workload.

It is recommended that for the load balance purpose, a proper dragonfly configuration should follow $a=2p=2h$ \cite{kim-micro}. Under this configuration, the total number of groups ``g" in the dragonfly network would be $g = a*h+1 $, the total number of compute nodes ``N" in the network would be $N = p*a*g $. An implementation of dragonfly network follows such configuration is illustrated in Figure \ref{fig:dragonfly-overview}. There are 6 routers in each group, each router has 3 compute nodes attached to it. This dragonfly implementation consists of 19 groups and 342 nodes in total.

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.48\textwidth]{dragonfly-overview}
  \caption{Dragonfly Network Overview. There are 19 groups, 6 routers in each group, 3 compute nodes connected to each router, 342 nodes in total. Job $J1$ requires for 7 nodes, gets its allocation by random placement. Job $J2$ requires for 8 nodes, gets its allocation by contiguous placement. }
  \label{fig:dragonfly-overview}
\end{figure}

@@

\textbf{Contiguous Placement:} In this policy, the compute nodes are assigned to job in a consecutive manner. The assignment fills up a group first, only crosses group boundaries and start to fill the next group as necessary. Shown in Figure \ref{fig:dragonfly-overview}, $J2$ gets the required amount of nodes by contiguous placement. Contiguous placement confines the tasks of a job into the same group, which may result in local network congestion and increase the possibility of hot-spots.

\textbf{Random Placement:} In this policy, job will get a set of nodes that randomly selected from all the available nodes in the system. The nodes assigned to $J1$ in Figure \ref{fig:dragonfly-overview} would be attached to different routers in different groups. Routers may be shared by different jobs when random placement is in use. Random placement can distribute the tasks of a job uniformly across the whole system, thus avoid the possible local congestion and reach load balance. However, random placement may cause possible congestion on global links.

\subsection{Routing on Dragonfly}
\label{sec:routing-schemes}
The routing policy refers to the strategy adopted to route a message (a stream of packets) from the source router to the destination router. Previously studied routing policies for dragonfly network include minimal routing, adaptive routing \cite{dally-dragonfly}, progressive adaptive routing\cite{jiang} and some of their variations \cite{won-prog-adaptive}. In this work we study three widely adopted routing policies on dragonfly networks.

\textbf{Minimal:} In this policy, a message from source router takes the the shortest path to the destination router. When there are multiple shortest paths between the source and destination router, the message will be evenly divided among the paths. Minimal routing can guarantee the minimum hops the message takes from source to destination. However, it may result in congestion along the shortest path.

\textbf{Adaptive:} In this policy, the path a message takes will be adaptively chosen between shortest and non-shortest path, depending on the congestion situation along that path. For the non-shortest path, an intermediate router will be randomly chosen. The message first takes the shortest path from source to that intermediate router, then it takes the shortest path from that intermediate router to the destination. Since routing path is chosen adaptively based on the congestion situation, adaptive routing can avoid local congestion and possible hot-spots. Adaptive routing collapses to minimal routing in the absence of traffic congestion.

\textbf{Progressive Adaptive:} In this policy, the decision to route minimally at each hop in the source group will be re-evaluated. Any decision to route non-minimally at source router or at a subsequent hop is permanent and won't be re-evaluated \cite{jiang}. The progressive adaptive is capable of handling the case where a global link is congested but this information has not yet propagated back to the source router \cite{jiang}. The packet will take the non-minimal route only when the congestion is encountered.

@@

It is difficult to experiment with concurrently running jobs on HPC system \cite{zhou-ipdps-2015}\cite{jain-sc14}\cite{bhatele-sc11}\cite{jokanovic-ipdps-2015}. One reason is that job placement and routing policies are part of system configuration, it is impossible for user to make changes at will. Another reason is that it is unrealistic to reserve the system exclusively to run the same batch of jobs with desired placement and routing configurations and compare the results. Last but not least, configurable dragonfly networks that allow us to perform the exploration presented in this paper are hard to come by for the time being. Therefore, we resort to simulation in our work.

\subsection{Simulation Tool}
\label{sec:simulation-tool}
A simulation toolkit named CODES, Co-Design of Multilayer Exascale Storage Architectures, enables the exploration of simulating different HPC networks with high fidelity and great scalability \cite{codes}. CODES supports dragonfly network\cite{codes-dragonfly} \cite{misbah-tpds}, as well as other topologies such as torus\cite{misbah-pads-2014}\cite{ning-pads-2011} and fat-tree\cite{ning-pads-2015}, with packet-level high-fidelity simulation. It can drive these models through an MPI simulation later that utilizing DUMPI traces\cite{sst}. In this work, we perform an in-depth analysis of network performance as well as job interference when different job placement and routing policies are in use on dragonfly network.

@@

The parameters for building the dragonfly network been studied in our work are inspired the model proposed by Kim et. al \cite{kim-micro}. Our dragonfly network consists of 33 groups, each group contains 8 routers, and each router has 4 compute nodes attached to it. Overall, there are 264 routers and 1056 compute nodes in the network. The aggregated bandwidth of terminal link,  local and global channels are proportional to those of Cray Cascade system\cite{faanes}. In this work, we study network performance as well as job interference when 6 different job placement and routing policy combinations are in use on the dragonfly network, which are summarized in Table \ref{tab: placement routing configs}.\footnote{With respect to random placement, we experiment with 50 sets of distinctive allocation generated by random placement. The corresponding experimental results are median chosen, which intended to eliminate the possibility of variation.}

@@

    \item \textbf{Network Traffic:} The traffic refers to the amount of data go through each port of the router. For each router, the traffic go through its terminal link, local channel and global channel are analyzed respectively.

    \item \textbf{Network Saturated Time:} The saturated time refers to the time period when the buffer of certain port is full. Saturated time can denote the congestion on certain port.

@@

Since our study focus on network performance and job interference during communication, the time spent in computation by each MPI rank will not be counted in our experiments.

@@

Two sets of parallel workloads are used in this work. Workload \Rmnum{1} consists of AMG, MultiGrid and CrystalRouter. As we see in Table \ref{tab:apps-detail}, AMG has the least amount of data transfer, making it the least communication-intensive job in Workload \Rmnum{1}. Workload \Rmnum{2} consists of sAMG, MultiGrid and CrystalRouter. sAMG, a synthetic version of AMG, is generated by increasing AMG data transfer amount by 100x. sAMG is the most communication-intensive job in Workload \Rmnum{2}.

@@

%Table \ref{tab:apps-detail} summarizes the details of each application. It shows the number of MPI ranks, the average data transfer amount of each rank and the total data transfer amount for each application.

@@

\section{Study of Parallel Workload \Rmnum{1}}
%\section{Identify the ``Bully"}
\label{sec:workload-1}

We study the network performance as well as job interference from two perspectives. First, we analyze the network performance when Workload \Rmnum{1} is running under different placement and routing configurations. Then, we isolate each job from the workload, and analyze the interference by scrutinizing the traffic go through the routers that serving each job. These analysis allow us to identify the ``bully" in the workload.

@@

The average communication time spent by all MPI ranks when Workload \Rmnum{1} is running under different placement and routing configurations shown in Table \ref{tab:wkld-commtime}. The random placement policy performs comparably with all routing policies and outperforms all contiguous placement configurations. The contiguous placement coupled with minimal routing (CM) results in highest communication time. When coupled with (progressive) adaptive routing (CA, CPA), the communication time can be reduced but still higher than random placement polices.

@@

    \centering   
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/gc-stime}
        \caption{Global Channel Saturated Time}
        \label{fig:global-channel-stime}
    \end{subfigure}\hfill
     \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/lc-stime}
        \caption{Local Channel Saturated Time}
        \label{fig:local-channel-stime}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/tl-stime}
        \caption{Terminal Link Saturated Time}
        \label{fig:terminal-link-stime}
    \end{subfigure}%
   \caption{The aggregate traffic and saturated time of the dragonfly network when Workload \Rmnum{1} running under 6 different placement and routing configurations. ``CA" and ``CPA" performs comparably, the corresponding lines are overlapped. }
   \label{fig:wkld-network-traffic-stime}
\end{figure*}

Figure \ref{fig:wkld-network-traffic-stime} shows the aggregate traffic for terminal links, local channels and global channels as well as the corresponding saturated time when Workload \Rmnum{1} is running on dragonfly network with 6 different placement and routing configurations. When contiguous placement and minimal routing(CM) is in use, all the traffic goes within the contiguously allocated groups, causing congestion along certain shortest paths from source to destination routers. Both the traffic and the corresponding saturated time of global channels are the highest compared with other configurations, shown as the blue dashed line in Figure \ref{fig:global-channel-traffic} and \ref{fig:global-channel-stime}.
When contiguous placement is coupled with (progressive) adaptive routing (CA and CPA), the traffic through congested global channels will be greatly reduced, so does the corresponding saturated time, shown as the red and green dashed line in Figure \ref{fig:global-channel-traffic} and \ref{fig:global-channel-stime}. Since (progressive) adaptive routing are congestion aware, the traffic would be redirected to the intermediate routers in other idle groups when congestion occurs, thus the traffic burden on the contiguously allocated groups will be alleviated.

Random placement policy performs comparably with three different routing policies.
Shown as the solid lines in Figure \ref{fig:global-channel-traffic}, the traffic through global channels are quite balanced across all the routers, thus no particular global channel will suffer high traffic burden. When the random placements coupled with minimal routing (RM), the packets can avoid unnecessary intermediate forwarding as introduced by (progressive) adaptive routing, thus generating less traffic. That's why the blue solid line is lower than the red and green solid line. Since the random placement can uniformly distribute traffic, there are no sudden surge of saturated time of global channel over the network, shown as solid lines in Figure \ref{fig:global-channel-stime}.

@@

We observe similar phenomena from the local channel traffic and saturated time, shown in Figure \ref{fig:local-channel-traffic}, \ref{fig:local-channel-stime}). Contiguous placement coupled with minimal routing (CM) results in local congestion, and thus the aggregate traffic through local channels of certain routers is quite high. Even with adaptive or progressive adaptive routing, the local congestion caused by grouping all MPI ranks in the same group(s) in contiguous placement can not be eliminated. Random placement is effective at addressing this type of congestion.

%terminal link traffic and busy time explanation
Figure \ref{fig:terminal-link-traffic} is for the purpose of symmetry, it shows the traffic per rank (aka traffic through terminal links), which stay the same when different placement and routing configurations are in use. All 6 lines are overlapped in Figure \ref{fig:terminal-link-traffic}. However, due to the traffic changes on global and local channels, the saturated time of terminal links will experience variability when different routing schemes are in use, as shown in Figure \ref{fig:terminal-link-stime}. When contiguous placement coupled with minimal routing(CM), the accumulated congestion on local and global channels cause extremely long saturated time on the terminal link, shown as the blue dashed line in Figure \ref{fig:terminal-link-stime}. This again manifests the severe congestion caused by using contiguous placement and minimal routing.

@@

As shown in Figure \ref{fig:mg-commtime}, \ref{fig:cr-commtime}, MultiGrid and CrystalRouter suffer the longest communication time when running with contiguous placement and minimal routing (CM). When contiguous placement coupled with adaptive or progressive adaptive routing (CA, CPA), the communication time of both MultiGrid and CrystalRouter are significantly reduced. Random placement further reduces their communication time, and reaches the best performance when coupled with (progressive) adaptive routing (RPA, RA). Both MultiGrid and CrystalRouter show the same trend when they are running under different placement and routing configurations. AMG is quite an exception. The communication time of AMG skyrockets when running under random placement coupled with (progressive) adaptive routing (RPA, RA), shown in Figure \ref{fig:amg-commtime}. The contiguous placement coupled with (progressive) adaptive routing can reach the best performance for AMG. CM is no better than RM, but still much better RPA, RA.

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/amg/lc-traffic}
        \caption{AMG Local Channel Traffic}
        \label{fig:amg-lc-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/mg/lc-traffic}
        \caption{MG Local Channel Traffic}
        \label{fig:mg-lc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/cr/lc-traffic}
        \caption{CR Local Channel Traffic}
        \label{fig:cr-lc-traffic}
    \end{subfigure}\\
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/amg/gc-traffic}
        \caption{AMG Global Channel Traffic}
        \label{fig:amg-gc-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/mg/gc-traffic}
        \caption{MG Global Channel Traffic}
        \label{fig:mg-gc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/cr/gc-traffic}
        \caption{CR Global Channel Traffic}
        \label{fig:cr-gc-traffic}
    \end{subfigure}%
   \caption{The aggregate traffic go through the local and global channels of routers serving specific application. ``CA" and ``CPA" perform comparably, the corresponding lines are overlapped. More routers are involved in serving each application when random placement is in use, compared with contiguous placement.}
   \label{fig:3app-lc-gc-traffic}
\end{figure*}

@@

\emph{The performance of less communication-intensive jobs in workload will suffer degradation when network random placement and adaptive routing are in use.} AMG in Workload \Rmnum{1} are ``bullied" by their concurrently running communication-intensive peers, MultiGrid and CrystalRouter. AMG shares routers and groups with MultiGrid and CrystalRouter when random placement is in use. The traffic from MultiGrid and CrystalRouter will be redirected to the routers that serving AMG, slowing down AMG's communication. \footnote{We have tried three different congestion sensing schemes that used in adaptive routing\cite{won-prog-adaptive}, although there are some variations between the results, none of the congestion sensing scheme can prevent the ``bully" from happening.}

\emph{Consistent performance of each application can be guaranteed when contiguous placement and minimal routing are in use.} The router and group sharing between jobs are prohibited in contiguous placement policy. The traffic won't be redirected from congested routers to less busy ones when minimal routing is in use. Thus the interference between concurrently running jobs will be eliminated and the ``bully" behavior will be terminated. 
%However, contiguous placement and minimal routing usually results in local congestion and unbalanced network utilization.

@@

In this section, we will verify the observations made in the previous section when Workload \Rmnum{1} is running on the dragonfly network under different placement and routing configurations. We conduct the same sets of experiments with Workload \Rmnum{2}, which consists of sAMG, MultiGrid and CrystalRouter. By replacing AMG with sAMG, we turn the ``bully" into ``bullee".

@@

The average communication time spent by all MPI ranks when Workload \Rmnum{2} is running under different placement and routing configurations shown in Table \ref{tab:syn-wkld-commtime}. The contiguous placement coupled with minimal routing results in highest communication time. The random placement coupled with minimal routing is the most efficient. Progressive adaptive routing is better than adaptive routing when random placement is in use. And they have comparable results when coupled contiguous placement. The performance of Workload \Rmnum{2} under different configurations keep the similar pattern as it of Workload \Rmnum{1}.
